Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-02-12 Manuscript/manuscript.Rmd
manuscript edits

Identifying the mechanisms underlying the formation of TADs are a complex and active area of research [??? Need all refs that prove it]. The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes prompt the use of computational approaches to identify which epigenomic features are most predictive of the TAD boundaries. Current methods rely on the use of classification algorithms to draw associations between the spatial relationship of functional genomic elements and regions that include TAD boundaries [@mourad2016computational; @hong2017computational; @huang2015predicting]. Although several useful insights have been made, such as the convergent binding of CTCF [@ong2014ctcf; @ghirlando2016ctcf], many methods ignore key characteristics of domain data, such as heavily imbalanced number of boundaries vs. other regions. In a classification setting, when one is attempting to discriminate between regions with and without TAD boundaries, the class imbalance issue will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. Such heavily imbalanced datasets are known to adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence;@jeni2013facing]. Consequently, careful consideration must be taken when deciding which performance metric(s) to use for evaluation of predictive models, especially in the case of imbalanced data.

Commonly used measures of model performance for binary classifiers include the area under the receiver operating curve (AUROC). The popularity surrounding the use of ROC curves is that they do not depend on a threshold. However, when the prevalence of an event is low--that is, a defined genomic region containing a TAD boundary--mean probabilities are biased towards the majority class. This results in the area of interest under an ROC curve to be compressed toward a small corner in the ROC space [@prati2011survey]. While not directly affecting the value of the AUROC, this can have severe consequences on the interpretability of the model's performance. A better threshold-free measure to use when dealing with imbalanced data in the area under the precision-recall curve (AUPRC). It has been shown that the use of AUPRC with imbalanced datasets yields more accurate and intuitive interpretation of model performance when compared to AUROC [@saito2015precision]. Likewise, metrics like accuracy and specificity are also problematic when evaluating model performance on imbalance data due to being highly influenced by the correct classification of the majority class. Better measures to consider are the F1-Score and Matthew's Correlation Coefficient (MCC) [@jeni2013facing]. The F1-Score can be interpreted as a weighted average of the precision and recall values, while MCC is, in essence, the correlation coefficient between the observed and predicted binary classifications. Both metrics use all 4 quadrants of the confusion matrix, thus providing more balanced measures in the case of imbalanced data [@chicco2017ten].

>>>>

The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes prompt the use of computational approaches to identify which epigenomic features are most predictive of TAD boundaries. Current methods rely on the use of classification algorithms to draw associations between the spatial relationship of functional genomic elements and regions that include TAD boundaries [@mourad2016computational; @hong2017computational; @huang2015predicting]. Although several useful insights have been made, such as the convergent binding of CTCF [@ong2014ctcf; @ghirlando2016ctcf], many methods ignore key characteristics of domain data, such as the impact of heavily imbalanced number of boundaries vs. other regions. In a classification setting, when one is attempting to discriminate between regions with and without TAD boundaries, the class imbalance issue will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. Such heavily imbalanced datasets are known to adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence;@jeni2013facing]. Consequently, careful consideration must be taken when deciding which performance metric(s) to use for evaluation of predictive models, especially in the case of imbalanced data.

Commonly used measures of model performance for binary classifiers include the area under the receiver operating curve (AUROC). The popularity surrounding the use of ROC curves is that they do not depend on a threshold. However, when the prevalence of an event is low--that is, a defined genomic region containing a TAD boundary--mean probabilities are biased towards the majority class. This results in the area of interest under an ROC curve to be compressed toward a small corner in the ROC space [@prati2011survey]. While not directly affecting the value of the AUROC, this can have severe consequences on the interpretability of the model's performance. A better threshold-free measure to use when dealing with imbalanced data is the area under the precision-recall curve (AUPRC). It has been shown that the use of AUPRC with imbalanced datasets yields more accurate and intuitive interpretation of model performance when compared to AUROC [@saito2015precision]. Likewise, metrics like accuracy and specificity are also problematic when evaluating model performance on imbalance data due to being highly influenced by the correct classification of the majority class. Better measures to consider are the precision and recall themselves. Both the precision and recall of a model include the true positives in the numerator of their respective ratios, thereby yielding more appropriate measures of a model's ability to accurately classify positive classes. Therefore, certain composite metrics like the F1-Score, Youden's Index, and Matthew's Correlation Coefficient (MCC) offer better interpretation, especially in the instance of imbalanced classes [@jeni2013facing]. The F1-Score can be interpreted as a weighted average of the precision and recall values, while Youden's index is interpreted as the arithmetic mean between sensitivity (recall) and specificity. MCC is, in essence, the correlation coefficient between the observed and predicted binary classifications. Each metric uses all 4 quadrants of the confusion matrix, thus providing more balanced measures in the case of imbalanced data [@chicco2017ten].
